---
title: "Sparklyr demo"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Motivation

- In this demo 
  1. we will install 
    - sparklyr
    - Apache spark (through sparklyr)
  2. Run some basic codes
  
Codes are pretty similar to what we did for `dbplyr`. 

The main difference is we can run the model on the cluster.

## Installation


### Installation of `sparklyr`
```{r}
system('java -version') # the default version of java on this server matches
#install.packages("sparklyr") # install package
```

### Install and connect to Apache Spark

```{r}
library(sparklyr)
library(tidyverse)
spark_install("2.3")
spark_installed_versions()

## Connect
sc <- spark_connect(master = "local", version = "2.3")
```

### Put the data on spark

```{r}
spark_covid <- spark_read_csv(
  sc, 
  "data/covid-data-all.csv.gz",
  escape = '"',
  memory = T)

head(spark_covid)
glimpse(spark_covid)

spark_covid_latest <- spark_covid %>% filter(date == "2020-08-08") 
```

### create data partition

For data partition, we use `sdf_random_split()` 

```{r}
data_splits <- 
  spark_covid_latest %>%
  sdf_random_split(training = .8, testing = .2,
                                seed = 42)
spark_covid_train <- data_splits$training
spark_covid_test <- data_splits$testing
```


### Visualisation 

- You can use ggplot after bringing data to R. But this could be quite costly. You can use `dbplot` package for a limited types of plotting.

```{r}
install.packages("dbplot")
library(dbplot)

dbplot_histogram(spark_covid_test, deaths_per1000) + 
  theme_minimal()
```

### Run models and evaluate

- Now we work with Sparks `MLlib` through `ml_*` function

#### Model fitting

```{r}
install.packages("tictoc")
tictoc::tic()
model_lm <- spark_covid_train %>% 
  ml_linear_regression(deaths_per1000 ~ Black + Hispanic + pct + Unemployment)
tictoc::toc()
ml_evaluate(model_lm, spark_covid_train)$mean_squared_error
```

#### Calculate MSE

```{r}

ml_predict(model_ml, spark_covid_test) %>%
mutate(resid = deaths_per1000 - prediction) %>%
summarize(mse = mean(resid ^ 2)) %>%
collect()

# or simply

ml_evaluate(model_lm, spark_covid_test)$mean_squared_error

```


### Example of other methods

- `ml_logistic_regression()`
- `ml_random_forest()`
- `ml_gradient_boosted_trees()`
